---
title: "FML_Clustering"
output:
  pdf_document: default
  html_document: default
date: "2023-11-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#We need the following packages to use the clustering needed for this assignment.

```{r}
library(flexclust)
library(cluster)
library(tidyverse)
```

```{r}

library(factoextra)
library(FactoMineR)
library(ggcorrplot)

```

```{r}
df_pharma <- read.csv("C:/Users/navat/Downloads/Pharmaceuticals.csv")

```

```{r}
summary(df_pharma)
head(df_pharma)

```
Q1) Use only the numerical variables (1 to 9) to cluster the 21 firms. Justify the various choices made in conducting the cluster analysis, such as weights for different variables, the specific clustering algorithm(s) used, the number of clusters formed, and so on. 


Sol: As mentioned, lets only select the variables from 1 to 9.

```{r}

df_pharma_1 <- df_pharma[3:11]
head(df_pharma_1)
summary(df_pharma_1)

```


#Let's normalise the data. Using the k-means clustering algorithm, this scales the data which will ensure that all the variables contributes equally and also calculating the distance. For the optimal number of clusters, we have used the fviz_nbclust with the "silhouette" method, which calculates the silhouette scores for different clusters and helps us to identify the number to maximise the seperation between clusters. 

```{r}
df_pharma_2 <- scale(df_pharma_1)
row.names(df_pharma_2) <- df_pharma[,1]
dist <- get_dist(df_pharma_2)
corr <- cor(df_pharma_2)
fviz_nbclust(df_pharma_2,kmeans, method = "silhouette")
```


#After the application, we can find that the optimal number of clusters to be 5. So, using k = 5 and number of restarts are 25.

```{r}
set.seed(69)
k_5 <- kmeans(df_pharma_2,centers = 5,nstart = 25)
print(k_5)
```

```{r}

k_5$centers
k_5$size

```

#We see that the within cluster sum of square value is 65.5% when k=5. The "centers" produced from the kmeans function provides us with the mean values of variables within clusters. 

#When k=5, using the fviz_cluster to form the clusters.
```{r}
fviz_cluster(k_5,data = df_pharma_2)
```

#Using the elbow method

```{r}
fviz_nbclust(df_pharma_2,kmeans,method = "wss")
```


#Using the "euclidean" distance
```{r}
dist <- dist(df_pharma_2,method="euclidean")
fviz_dist(dist)
```

#Using the "Manhattan" distance
```{r}
set.seed(69)

k_5.1 = kcca(df_pharma_2,k=5,kccaFamily("kmedians"))
k_5.1
```

```{r}
Cluster_in <- predict(k_5.1)
dist(k_5.1@centers)

image(k_5.1)
points(df_pharma_2,col=Cluster_in,pch=19,cex=0.3)

```

Q2) Interpret the clusters with respect to the numerical variables used in forming the clusters. Is there a pattern in the clusters with respect to the numerical variables (10 to 12)?

Sol: Here, we can see some patterns between the non-numeric variables to the numeric variable based clusters. 
For,
#Cluster1: These stocks has higher market cap, ROE and ROA comparing to other clusters but low leverage.
#Cluster2: These stocks are moderate in market cap, ROE, ROA and leverage.
#Cluster3: These stocks are moderate again where the recommendations are more halfway, not complete.
#Cluster4: The summary of cluster 4 is more in negatives, with second lowest market cap, ROE, ROA, lowest PE_ratio, Asset_turnover but highest Revenue growth.
#Cluster5: These stocks have the highest leverage but lowest market cap.


#Finding out the pattern between the numerical variables to non-numeric (10 to 12)
```{r}
df_pharma_1%>%mutate(Cluster=k_5$cluster)%>%group_by(Cluster)%>%
summarise_all("mean")

clusplot(df_pharma_2,k_5$cluster,main="Clusters",color = TRUE, labels = 3,lines = 0)

```

#Median recommendation for non-numeric value
```{r}
df_pharma_3 <- df_pharma[12:14] %>% mutate(Clusters=k_5$cluster)
ggplot(df_pharma_3, mapping = aes(factor(Clusters), fill 
=Median_Recommendation))+geom_bar(position='dodge')+labs(x ='Clusters')

```

#Location for non-numeric value
```{r}
ggplot(df_pharma_3, mapping = aes(factor(Clusters),fill =
Location))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```

#Exchange for non-numeric value 
```{r}
ggplot(df_pharma_3, mapping = aes(factor(Clusters),fill =
Exchange))+geom_bar(position = 'dodge')+labs(x ='Clusters')
```

#By looking at all the graphs, we can interpret the following:

#Cluster1: It has a hold and a moderate buy recommendation, situated in UK and US and are traded on NYSE.
#Cluster2: It is the only cluster with all the recommendations, in US, UK and Switzerland and are traded on NYSE.
#Cluster3: It is similar to cluster 1 in recommendation but are less in count, functions in canada and US, and are traded on NYSE likewise clusters 1-4.
#Cluster4: It is moderate in buy and sell for recommendations, functions in France, germany and US.
#Cluster 5: It is the only cluster which is traded on all the markets(AMEX,NASDAQ,NYSE) and are present in Germany and US.



Q3) Provide an appropriate name for each cluster using any or all of the variables in the dataset.

Sol:

#For,
#Cluster1: Topper with high scores since this has high market valuation, profit and low leverage.
#Cluster2: Medium scaler since these are moderate in all the sections.
#Cluster3: Moderate growth, since they are moderate in all the sections likewise cluster 2.
#Cluster4: Risk bull, since it has high risk and also higher revenue growth.
#Cluster5: Big market hero, since they are in all the exchange markets.





